---
title: "ISCD Homework 4"
author: "Anurag Bhunia(bmat2508@isibang.ac.in)"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridisLite)
library(ggplot2)
```

**Question 1**

a)

```{r}
set.seed(2026)
#Generating p
p=runif(1)
#Getting 100 Uniform(0,1) Samples
u=runif(100)
#Appling the inverse CDF transform
x=ceiling(log(1-u)/log(1-p))
x
set.seed(2025)
#Setting r=5
r=5
nb=numeric(100)
for(i in 1:100){
    U=runif(r)
    X=ceiling(log(1-U)/log(1-p))
    nb[i]=sum(X)
}
nb
```

b)

```{r}
set.seed(2023)
u=runif(100)
#Generating 100 Cauchy samples
c=tan(pi*(u-0.5))
c
```

2.

```{r}
set.seed(2024)
ex=rexp(100)
u=runif(100)
#Deviation of the exponential distribution
qqnorm(ex)
qqline(ex)
#Deviation of the uniform distribution
qqnorm(u)
qqline(u)
```

3.

0) This is known as the method of moments. Given a sample, we compute its mean and variance. Knowing the distribution, we equate both to their theoretical formulas (in this case, a binomial distribution). The expressions for p-hat and n-hat are exactly those which one would get after solving for the theoretical values of p and n respectively.
This is used to estimate the parameters pertaining to the given sample.

a)

```{r}
n=100
N=20
p=0.4
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
```

b) For n=1000, we have

```{r}
n=1000
N=20
p=0.4
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
```

The values are quite different from when we had just 100 samples.

c)

```{r}
#Bin(10,0.1)
n=100
N=10
p=0.1
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
n=1000
N=10
p=0.1
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
#Bin(10,0.9)
n=100
N=10
p=0.9
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
n=1000
N=10
p=0.9
x=rbinom(n,N,p)
m1=mean(x)
m2=mean(x^2)
Nhat=m1^2/(m1-(m2-m1^2))
phat=(m1-(m2-m1^2))/m1
#Estimated sample size
Nhat
#Estimated probability
phat
```

4.

a)

```{r}
x=seq(-2,2,by=0.1)
y=pnorm(x,0,1)
plot(x,y,type="l",main="CDF of Standard Normal Distribution")
```

b)

```{r}
m=100
p=0.4
x=seq(-2,2,by=0.1)
mu=m*p
sd=sqrt(m*p*(1-p))
c=x*sd+mu
y=pbinom(c,n,p)
plot(x,y,type="l",main="Standardized Binomial Distribution")
```

c)

```{r}
set.seed(1011)
m=100
p=0.4
Y=rbinom(1000,m,p)
#Histogram for Binomial Distribution
hist(Y)
mu=m*p
sd=sqrt(m*p*(1-p))
Z=rnorm(1000,mu,sd)
d=density(Z)
plot(d,type="l",main="Normal Density Plot")
```

From the 2 plots, one can conclude that the binomial histogram has a similarity in shape with the normal density plot. This is a consequence of the Central Limit Theorem, as the sample size increases, the binomial distribution ressembles the normal.

5.

```{r}
# Define the possible outcomes of a single die roll (1 through 6)
x = c(1,2,3,4,5,6)

# Define the probability for each outcome (1/6 for a fair die)
probx = c(1/6, 1/6, 1/6, 1/6, 1/6, 1/6)

# Simulate 1500 independent die rolls using the defined probabilities
Rolls = sample(x, size = 1500, replace = T, prob = probx)

# Organize the 1500 rolls into a matrix with 5 rows 
# This effectively creates 300 groups, where each group has 5 die rolls
Rollm = matrix(Rolls, nrow = 5)

# Calculate the sum of each column (the sum of the 5 rolls in each group)
Rollsums = apply(Rollm, 2, sum)

# Load the ggplot2 library for advanced data visualization
library(ggplot2)

# Manually define the Normal density function formula: 
# 1/(sqrt(2*pi)*s) * exp(-(x-a)^2 / (2*s^2))
density = function(x, a, s) { (1/((2*pi)^(0.5)*s )) * exp(-(x-a)^2/(2*s^2)) }

# Convert the vector of sums into a data frame format required by ggplot2
dfrolls = data.frame(Rollsums)

# Calculate the sample mean (mu) and standard deviation (sigma) of the sums
mu = mean(dfrolls$Rollsums)
sigma = sd(dfrolls$Rollsums)

# Create the visualization
ggplot(data = dfrolls) +
  # Draw a histogram of the sums; y=..density.. scales it to match a probability curve
  geom_histogram(mapping = aes(x = Rollsums, y = ..density..), color = "#00846b", fill = NA, binwidth = 1) +
  # Set the x-axis limits to focus on the range of possible sums (5 to 30)
  xlim(5,30) +
  # Overlay the manually defined Normal density function using the calculated mu and sigma
  geom_function(fun = density, args = list(a = mu, s = sigma), color = "black")
```

This is a demonstration of the Central Limit Theorem. It simulates rolling a fair die 1500 and recording sums of die rolls from sets of 5. The sum is distributed almost normally as its corresponding histogram aligns closely with the normal density curve.