---
title: "Intro to Statistics and Computation with Data"
author: "Anurag (bmat2508@isibang.ac.in)"
output: pdf_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(viridisLite)
library(UsingR)
library(MASS)
```
**Class 1, January 7, 2026**

Marks breakdown for this course:

- Homework: 5

- Quiz: 15

- Midterm: 20

- Worksheet: 10

- Final: 50

**Basic Syntax**
```{r}
#Creating a list of numbers
scores=c(21,23,34,54,12)
scores
mean(scores)
x=1:100
x[x>10|x<90]
```
There are many pre-installed datasets in R. 
```{r}
#Air quality data of New York City
head(airquality,9)
```
Here's how we can plot graphs based on given data:
```{r}
#Scatter plot of Ozone versus Wind
plot(airquality$Wind, airquality$Ozone,main="Ozone vs. Wind",xlab="Wind(mph)",ylab="Ozone(ppb)")
#What happens when you try to plot the whole dataset?
plot(airquality)
```

**Class 2, January 14 2026**
```{r}
#Using ggplot
ggplot(data=mpg)+geom_point(mapping=aes(x=displ,y=hwy))
#Adding more layers
ggplot(data=mpg)+geom_point(mapping=aes(x=displ,y=hwy,colour=class))+scale_colour_viridis_d(option="plasma")
#Using facet wrap
ggplot(data=mpg)+geom_point(mapping=aes(x=displ,y=hwy,colour=class))+scale_colour_viridis_d(option="plasma")+facet_wrap(~class,nrow=2)
#Using different plots
ggplot(data=mpg)+geom_bar(mapping=aes(x=class,fill=class))
#Summary
ggplot(data=mpg)+stat_summary(mapping=aes(x=class,y=hwy),fun.min=min,fun.max=max,fun=median)
#Adding more colour layers
ggplot(data=mpg)+geom_bar(mapping=aes(x=class,fill=trans))+scale_fill_viridis_d()
#Making a pie chart
ggplot(data=mpg)+geom_bar(mapping=aes(x=class,fill=class))+scale_fill_viridis_d()+coord_polar()
#Stacking bars
ggplot(data=mpg)+geom_bar(mapping=aes(x=class,fill=trans),position="dodge")+scale_fill_viridis_d()
#BoxPlot- cty for compact
mpgcs=subset(mpg,mpg$class=="compact")
ggplot(data=mpgcs,mapping=aes(x=class,y=cty))+geom_boxplot()+coord_flip()
#Jitter
ggplot(data=mpgcs,mapping=aes(x=class,y=hwy))+geom_boxplot()+geom_jitter(aes(colour=trans),size=0.4,alpha=0.9)+coord_flip()
mpgr=mpg[c("class","hwy","cty")]
mpgrc=subset(mpgr,mpgr$class=="compact")
rdf<-data.frame(Mileage=c(rep("hwy",nrow(mpgrc)),rep("cty",nrow(mpgrc))),compact=c(mpgrc$hwy,mpgrc$cty))
head(rdf)
tail(rdf)
ggplot(data=rdf,mapping=aes(x=Mileage,y=compact))+geom_boxplot()
```


Now we see how to simulate various probability distributions in R
```{r}
#Picking n many uniform random numbers 
sample(1:6,2)
#Rolling a dice ten times
sample(1:6,10,replace=T)
```


**Class 3, January 28 2026**

W would like to simulate various probability distributions in R. R can simulate both conitnuous and discrete uniform distributions as we saw in the previous class. 

```{r}
#Tossing a biased coin
sample(c(0,1),10,replace=T,prob=c(0.3,0.7))
#Simulating uniform random variables
set.seed(2021)
runif(3)
#Simulating Bernoulli Random Variables
p=0.5
U=runif(1,min=0,max=1)
X=as.numeric(U<p)
X
#Simulating Binomial Random Variables
#We want to similate a Bin(50,0.2) random variable.
p=0.2
U=runif(50,min=0,max=1)
X=as.numeric(sum(U)<10)
X
#Simulating a p1,p2,...,pn distribution
p1=0.35
p2=0.15
p3=0.4
p4=0.1
P=c(p1,p1+p2,p1+p2+p3,p1+p2+p3+p4)
X=c(1,2,3,4)
counter=1
set.seed(2021)
r=runif(1,min=0,max=1);
while(r>P[counter])
    counter=counter+1
end
X[counter]
#Simulating Exponential Random Variables
U=runif(1,min=0,max=1)
lambda=1
X=-(1/lambda)*log(U)
X
#Simulating Gamma Random Variables
n=20;
lambda=10;
X=(-1/lambda)*sum(log(runif(n,min=0,max=1)))
X
#Simulating Normal Random Variables (Box-Muller Transformations)
U1=runif(1)
U2=runif(1)
Z1=sqrt(-2*log(U1))*cos(2*pi*U2)
Z1=sqrt(-2*log(U1))*sin(2*pi*U2)
#Creating functions in R
f=function(x,y){
    x+y
}
f(2,2)
f=function(x){dbinom(x,5,0.3)}
f(0:5)
#Plotting the CDF of Normal Distribution
x=seq(-3,3,by=0.1)
y=pnorm(x)
plot(x,y,type="l")
#Getting the Q-Q PLot
x=rnorm(1000)
qqnorm(x)
qqline(x)
```

```{r}
x=seq(-3,3,by=1/1000)
densityx=dnorm(x)
plot(x,densityx,type="l",col="red")
x=rnorm(100)
y=rt(100,30)
par(mfrow=c(1,3))
boxplot(x,y)
qqnorm(x);qqline(x)
qqnorm(y);qqline(y)
#Binomial CLT
N=1000
n=100
p=0.99
x=rbinom(N,n,p)
Z=(x-n*p)/sqrt(n*p*(1-p))
hist(Z)
qqnorm(Z)
qqline(Z)
#Using package MASS
par(mfrow=c(1,1))
plot(Animals$body,Animals$brain)
cor(rank(Animals$body),rank(Animals$brain))






```


