---
title: "ISCD Worksheet 5"
author: "Anurag Bhunia(bmat2508@isibang.ac.in)"
output: pdf_document
---

1. 

```{r}
x=c(3,4,2,5,2,4,4,6,3,4)
y=ecdf(x)
plot(ecdf(x),main="Empirical Distribution")
```

2.

```{r}
replicate(10, {
  u=runif(10000)
  v=runif(10000)
  z=sqrt(u^2+v^2)
  sum(z<1)/10000
})
```
 This code takes 10000 points from the unit interval [0,1] once each for u and v. Then it computes the proportion of those points such that (u,v) is less than a unit distance apart from the origin. It is equivalent to solving an integral which comes out to be pi/4. In this case as well, that proportion is quite close to the theoretical value. 
 Yes, the code can be modified, just multiplying the proportion by 4 would give us an approximate value of pi.

3.

```{r}
L=function(p){
    N=45
    x=26
    choose(N,x)*p^x*(1-p)^(N-x)
}
plot(L,from=0,to=1)
optimize(L,interval=c(0,1),maximum=TRUE)
```

This code is first creating a function to compute the probability that a binomial random variable with parameters 45 and p would take the value 26. Then it plots the function from 0 to 1. Finally, the optimize bit of the code is showing the value of p at which the likelihood gets maximised and it also returns the maximum value of that likelihood.
